---
title: "Analysis of Subfamili"
author: "Robin von Allmen"
date: "18 3 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
require(Morpho)
require(rgl)
options(rgl.useNULL = TRUE) # Suppress the separate window.
knitr::knit_hooks$set(webgl = hook_webgl)
library(geomorph)
setwd("~/GitHub/Ichneumonidae/R-Scripts/Determine Higher-Grouping")
```

## Preparing Data

Path to Data: "~/GitHub/Ichneumonidae/R-Scripts/Determine Higher-Crouping"

Data contains 234 specimens from Ophioniformes/Ichneumoniformes/Pimpliformes. However, we will only select certain subfamilies (Banchinae/Ophioninae/Pimplinae/Ichneumoninae/Phygadeuontinae (here still marked as Gelinae from Townes 1969)).

!!! This is the first time these subfamilies have been compared with geometric morphometrics by forewing venation!!!

After Data-Aquisation we superimpose the Data with the function gpagen().

```{r cars, include=FALSE, results= 'hide'}

landmark = readland.tps("blockcourse_highersub.tps",specID = "ID", readcurves = T, warnmsg = TRUE)

##      define classifiers (subfam or tribes)
species = as.factor( substr( dimnames(landmark)[[3]], 1, 6))
subfam = as.factor( substr( dimnames(landmark)[[3]], 8, 13))
areolet = as.factor( substr( dimnames(landmark)[[3]], 15, 15))
tribe = as.factor( substr( dimnames(landmark)[[3]], 17, 19))


##      define Semilandmarks, first LM is the start, last one the end (those are both
##      fixed landmarks, in between are the semilandmarks):
sliders = define.sliders(c(15,22:29,16))

##      procrustes fit - "superimposition" - rotates and scales the landmarks to their closest distance,
##      minimizing the shape and size difference:
proD<-gpagen(landmark, surfaces = NULL, curves= sliders,
             PrinAxes = FALSE, max.iter = NULL, ProcD = F, Proj = TRUE,
             print.progress = TRUE)
use= c("pimpli", "banchi", "phygad", "ophion", "fossil")
reduced<- proD$coords[,,subfam%in% use]
pro_red<-gpagen(reduced, surfaces = NULL, curves= sliders,
             PrinAxes = FALSE, max.iter = NULL, ProcD = F, Proj = TRUE,
             print.progress = TRUE)
subfam1<-as.factor(substr(dimnames(pro_red$coords)[[3]],8,13))
species1<-as.factor(substr(dimnames(pro_red$coords)[[3]],1,6))
tribe1<-as.factor(substr(dimnames(pro_red$coords)[[3]],17,19))
```

## PCA Analysis
A PCA was done with the subfamilies Pimplinae, Gelinae, Ophioninae, Ichneumoninae and Banichnae.The ophioninae are seperated from the rest of the subfamilies quite strongly.The pimplinae are also partially seperated. They partly overlap with Gelinae and Ichneumoninae. The Gelinae overlap mostly with Banchinae and partially with Pimplinae. Overall the subfamilies cluster mostly in groups. However, there are some outliers clustering within another group.   

```{r PCA, fig.show='hide',results='hide'}

PCA <-plotTangentSpace(pro_red$coords, groups=subfam1, legend=T)


```


```{r plot,echo=FALSE}
  ## the more beautiful version:

mat <- matrix(c(4,5,0,1,1,2,1,1,3), 3)
layout(mat, widths=c(1,1,1), heights=c(1,1,1))# set the size of the rows and columns

my_colors = matrix( NA, dim(PCA$pc.scores[,]))
for(i in 1:length( row.names(PCA$pc.scores[,]))){
my_colors[i] <- switch(substr(row.names(PCA$pc.scores[,])[i],8,13), "banchi" = "goldenrod1", "fossil"="black", "ophion" = "blueviolet",  "phygad" = "violet","pimpli" = "chartreuse3")
}

par(mar=c(4, 4, 1, 1)) # sets the margins
plot(PCA$pc.scores[,1],PCA$pc.scores[,2], col=my_colors, asp=T, las=1,pch=16, cex=1.5)
legend("topright", legend= levels(subfam1), col= c("goldenrod1", "black", "blueviolet", "violet", "chartreuse3"), pch=rep(19,14) )


segments(min(PCA$pc.scores[,1]), 0,  max(PCA$pc.scores[,1]),0, col="grey", lty="dashed")
segments(0, min(PCA$pc.scores[,2]), 0,  max(PCA$pc.scores[,2]), col="grey", lty="dashed")

# in order to name individuals: text(PCA$pc.scores[,1],PCA$pc.scores[,2], species)

ref <- mshape(proD$coords)

par(mar = c(0,0,0,0)) # sets the margins

plotRefToTarget(ref,PCA$pc.shapes$PC1min)
# Item 3
plotRefToTarget(ref,PCA$pc.shapes$PC1max)
# Item 4
plotRefToTarget(ref,PCA$pc.shapes$PC2max)
# Item 5
plotRefToTarget(ref,PCA$pc.shapes$PC2min)
```

## CVA
A CVA was done with the subfamilies Pimplinae, Gelinae, Ophioninae, Ichenumoninae and Banichnae.The Ophioninae seperate very strong from the rest of the subfamilies. The Gelinae are overlapping with the Ichneumininae. The Pimplinae and Banchinae are overlapping as well. The Pimplinae have a small overlap with the Ichneumoninae.Overall the subfamilies cluster mostly in groups. However, there are few outliers that cluster in another group.
The 95% confidence ellipses show the probablity of the dots really being in the ellpipses. Here, the probabilty is 95%.
```{r CVA}
#CVA
cva<- CVA(pro_red$coords, groups = subfam1, plot= T, rounds= 1000)

# plot the CVA

mat <- matrix(c(4,5,0,1,1,2,1,1,3), 3)
layout(mat, widths=c(1,1,1), heights=c(1,1,1))

#plot the data
par(mar=c(4, 4, 1, 1))
plot(cva$CVscores, col=my_colors, pch=16,asp=T,
     xlab=paste("1st canonical axis", paste(round(cva$Var[1,2],1),"%")),
     ylab=paste("2nd canonical axis", paste(round(cva$Var[2,2],1),"%")),
    las=1, cex=1)

##  add hulls to plot (merge groups)
for(jj in 1:length(levels(subfam1))){
  ii=levels(subfam1)[jj]
  kk=chull(cva$CVscores[subfam1==ii,1:2])
  lines(cva$CVscores[subfam1==ii,1][c(kk, kk[1])],
        cva$CVscores[subfam1==ii,2][c(kk, kk[1])], col=jj)

}

#lines that seperate the different types of wings
segments(min(cva$CVscores[,1]), 0,  max(cva$CVscores[,1]),0, col="grey", lty="dashed")
segments(0, min(cva$CVscores[,2]), 0,  max(cva$CVscores[,2]), col="grey", lty="dashed")

#links for the warpgrid that defines how LM are interlinked
array1 <- c(1,1,1,2,2,3,3,4,4,5,5,6,7,8,9,9,10,10,11,11,12,14,14,16,16,17,18,20)
array2 <- c(2,5,8,3,5,4,7,12,20,8,6,7,12,9,19,10,11,17,15,12,13,15,20,17,21,18,19,21)
warpgrid <- array(c(array1, array2), dim= c(28,2))



par(mar = c(0,0,0,0))



#plot the mean against the maxima of different shapes.
plotRefToTarget(ref,PCA$pc.shapes$PC1min, method = "points", links= warpgrid, col = "pink")
plotRefToTarget(ref,PCA$pc.shapes$PC1max, method = "points", links= warpgrid)
plotRefToTarget(ref,PCA$pc.shapes$PC2min, method = "points", links= warpgrid)
plotRefToTarget(ref,PCA$pc.shapes$PC2max, method = "points", links= warpgrid)

```

```{r Procrustes Distances, message = FALSE, results='hide', include=FALSE}
library(Evomorph)

inlist = list(pro_red$coords [,,])
for ( j in 1: dim(pro_red$coords )[3] )
  inlist[[j]] = pro_red$coords [,,j]


ID = paste (subfam1,species1, sep = "_", collapse = NULL, recycle0 = FALSE)

### PC1 for PC1dist thingy.. for empty matrix: ###################
PCA_sc<-PCA$pc.scores
row.names(PCA_sc) = ID
PCdist = as.matrix( dist(PCA_sc))

###  make empty matrix 
Proc_dist= matrix( NA, dim(PCA_sc)[1], dim(PCA_sc)[1] )

colnames(Proc_dist) = ID
rownames(Proc_dist) = ID

###  make a loop : calculate proc distances for each row. 
Proc_dist
for ( i in 1:length(ID)){
  Proc_dist[i,] = ShapeDist(inlist, inlist[[i]])
}

write.table(Proc_dist,"Proc_distances_ban_pim_oph_cte.txt", sep=" ",row.names = TRUE, col.names = TRUE)

###################################################################
Proc_dist = read.table("Proc_distances_ban_pim_oph_cte.txt", header=T)
```
## Cluster Analysis
```{r Clusters,}

data = data.frame(species1, subfam1, PCA$pc.scores[,1:25])

# Compute distances and hierarchical clustering

de <- dist(scale(data[c(3:27)]), method = "euclidean")

## Ward's minimum variance method

hd <- hclust(de, method = "ward.D2")

hd$labels = paste (subfam1,species1, sep = "_", collapse = NULL, recycle0 = FALSE)
# Convert hclust into a dendrogram and plot

hce <- as.dendrogram (hd)
```
## Dendrogramm
```{r Bäumli, results=FALSE, message=FALSE, fig.width= 9, fig.height= 20}
require("dendextend") ## needs a dendrogram

par(mar= c(10,3,1,5))
Label <- hce%>%labels
my_colors1 = matrix( NA, dim(Proc_dist))
for(i in 1:length(row.names(Proc_dist))){
my_colors1[i] <- switch(substr(Label[i],1,6), "banchi" = "cyan", "pimpli" = "yellow", "gelina" = "green", "ophion" = "darkblue", "ichneu" = "red")
}
hce %>%
  set("labels_col", value =  my_colors1) %>%
  set("branches_k_color", value = c("skyblue", "darkorange", "grey","darkorange4","deepskyblue4",
                                    "darkolivegreen3","darkorchid1","firebrick","gold1","aquamarine2","brown2","darkseagreen3","black"),k = 13) %>%
  set("labels_cex", 0.7)  %>%
  set("leaves_pch", 19)  %>% 
  set("nodes_cex", 0.4) %>%
  plot(axes=FALSE, main= "Dendrogram with ward.D2", horiz = TRUE)

# Add the colored bar
#colored_bars(colors = my_colors1, rowLabels = "subfam")
legend("topleft", legend=levels(subfam1), col= c("cyan", "green", "red", "darkblue", "yellow"), pch=15)
```

Principalcomponents 1:25, which explain about 98% of variation, were used to create this dendrogram with the hclust method "ward.D2". Even though there are outliers within this plot we clearly see, that our main higher groupings are clustered as we would expect it. 
