---
title: "First_Analysis"
author: "Robin von Allmen"
date: "18 3 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
require(rgl)
options(rgl.useNULL = TRUE) # Suppress the separate window.
knitr::knit_hooks$set(webgl = hook_webgl)
library(geomorph)
setwd("~/GitHub/Ichneumonidae/R-Scripts/Determine Higher-Crouping")
```

## Preparing Data

Path to Data: "~/GitHub/Ichneumonidae/R-Scripts/Determine Higher-Crouping"

Data contains 234 specimens from Ophioniformes/Ichneumoniformes/Pimpliformes. However, we will only select certain subfamilies (Banchinae/Ophioninae/Pimplinae/Ichneumoninae/Phygadeuontinae (here still marked as Gelinae from Townes 1969)).

!!! This is the first time these subfamilies have been compared with geometric morphometrics by forewing venation!!!

After Data-Aquisation we superimpose the Data with the function gpagen().

```{r cars, include=FALSE}

landmark = readland.tps("blockcourse_highersub.tps",specID = "ID", readcurves = T, warnmsg = TRUE)

##      define classifiers (subfam or tribes)
species = as.factor( substr( dimnames(landmark)[[3]], 1, 6))
subfam = as.factor( substr( dimnames(landmark)[[3]], 8, 13))
areolet = as.factor( substr( dimnames(landmark)[[3]], 15, 15))
tribe = as.factor( substr( dimnames(landmark)[[3]], 17, 19))


##      define Semilandmarks, first LM is the start, last one the end (those are both
##      fixed landmarks, in between are the semilandmarks):
sliders_red = define.sliders(c(14,21:28,15))

##      procrustes fit - "superimposition" - rotates and scales the landmarks to their closest distance,
##      minimizing the shape and size difference:
proD<-gpagen(landmark[2:29,,], surfaces = NULL, curves= sliders_red,
             PrinAxes = FALSE, max.iter = NULL, ProcD = F, Proj = TRUE,
             print.progress = TRUE)
use= c("pimpli", "banchi", "gelina", "ophion", "ichneu")
reduced<- proD$coords[,,subfam%in% use]
pro_red<-gpagen(reduced, surfaces = NULL, curves= sliders_red,
             PrinAxes = FALSE, max.iter = NULL, ProcD = F, Proj = TRUE,
             print.progress = TRUE)
subfam1<-as.factor(substr(dimnames(pro_red$coords)[[3]],8,13))
species1<-as.factor(substr(dimnames(pro_red$coords)[[3]],1,6))
tribe1<-as.factor(substr(dimnames(pro_red$coords)[[3]],17,19))
```

## PCA Analysis
Here, A PCA  with the subfamilies Pimplinae, Gelinae, Ophioninae, Ichenumoninae and Banichnae without landmark 1 was done. The ophioninae are seperated from the rest of the subfamilies quite strongly.The pimplinae are also partially seperated. They partly overlap with Gelinae and Ichneumoninae. The Gelinae overlap mostly with Banchinae and partially with Pimplinae. Overall the subfamilies cluster mostly in groups. However, there are some outliers clustering within another group. The plot doesn't really differ from the plot when you include landmark 1. 

```{r pressure, fig.show='hide', results='hide'}

PCA <-plotTangentSpace(pro_red$coords, groups=subfam1, legend=T)


```

One can work on this plot to get a beautiful shape so that it loocks somewhat like this:

```{r plot,echo=FALSE}
  ## the more beautiful version:

mat <- matrix(c(4,5,0,1,1,2,1,1,3), 3)
layout(mat, widths=c(1,1,1), heights=c(1,1,1))# set the size of the rows and columns

par(mar=c(4, 4, 1, 1)) # sets the margins
plot(PCA$pc.scores[,1],PCA$pc.scores[,2], col=subfam1, asp=T, las=1,pch=16, cex=1.5)
legend("topright", legend= levels(subfam1), col=palette(), pch=rep(19,14) )

segments(min(PCA$pc.scores[,1]), 0,  max(PCA$pc.scores[,1]),0, col="grey", lty="dashed")
segments(0, min(PCA$pc.scores[,2]), 0,  max(PCA$pc.scores[,2]), col="grey", lty="dashed")
# in order to name individuals: text(PCA$pc.scores[,1],PCA$pc.scores[,2], species)

ref <- mshape(pro_red$coords)
#qw= c(1,4,2,6,5,7,3)
par(mar = c(0,0,0,0)) # sets the margins

plotRefToTarget(ref,PCA$pc.shapes$PC1min)
# Item 3
plotRefToTarget(ref,PCA$pc.shapes$PC1max)
# Item 4
plotRefToTarget(ref,PCA$pc.shapes$PC2max)
# Item 5
plotRefToTarget(ref,PCA$pc.shapes$PC2min)
```


with the PCA scores u are also able to generate 3d plots which compare several Principal Components at once:

```{r 3D, webgl = TRUE}
colors = c("red","cyan","magenta","yellow","green")[subfam1]
plot3d(PCA$pc.scores[,1:3], col = colors, size = 9)
text3d(PCA$pc.scores[,1:3],texts=species1,pos=4,cex= 0.6)
legend3d("topright", legend = levels(subfam1), col = c("red","cyan","magenta","yellow","green") , pch= 16)
```

## Cluster Analysis
```{r Procrustes Distances, message = FALSE, results='hide', include=FALSE}
library(Evomorph)

inlist = list(pro_red$coords [,,])
for ( j in 1: dim(pro_red$coords )[3] )
  inlist[[j]] = pro_red$coords [,,j]


ID = paste (subfam1,species1, sep = "_", collapse = NULL, recycle0 = FALSE)

### PC1 for PC1dist thingy.. for empty matrix: ###################
PCA_sc<-PCA$pc.scores
row.names(PCA_sc) = ID
PCdist = as.matrix( dist(PCA_sc))

###  make empty matrix 
Proc_dist= matrix( NA, dim(PCA_sc)[1], dim(PCA_sc)[1] )

colnames(Proc_dist) = ID
rownames(Proc_dist) = ID

###  make a loop : calculate proc distances for each row. 
Proc_dist
for ( i in 1:length(ID)){
  Proc_dist[i,] = ShapeDist(inlist, inlist[[i]])
}

write.table(Proc_dist,"Proc_distances_ban_pim_oph_cte.txt", sep=" ",row.names = TRUE, col.names = TRUE)

###################################################################
Proc_dist = read.table("Proc_distances_ban_pim_oph_cte.txt", header=T)
```
## Cluster Analysis
```{r Clusters,}

data = data.frame(species1, subfam1, PCA$pc.scores[,1:25])

# Compute distances and hierarchical clustering

de <- dist(scale(data[c(3:27)]), method = "euclidean")

## Ward's minimum variance method

hd <- hclust(de, method = "ward.D2")

hd$labels = paste (subfam1,species1, sep = "_", collapse = NULL, recycle0 = FALSE)
# Convert hclust into a dendrogram and plot

hce <- as.dendrogram (hd)
```
## Dendrogramm
```{r Bäumli, results=FALSE, message=FALSE}
require("dendextend") ## needs a dendrogram

par(mar=c(10,3,1,1))
my_colors1 <- ifelse(substr(hd$labels,1,6)=="banchi", "cyan",ifelse(substr(hd$labels,1,6)=="gelina","green", ifelse(substr(hd$labels,1,6)=="ichneu", "red",ifelse(substr(hd$labels,1,6)=="ophion", "darkblue",ifelse(substr(hd$labels,8,13)== "fossil", "magenta", "yellow" )))))
hce %>%
  set("labels_col", value = c("skyblue", "darkorange", "grey","darkorange4","deepskyblue4",
                              "darkolivegreen3","darkorchid1","firebrick","gold1","aquamarine2","brown2","darkseagreen3","black"), k=13) %>%
  set("branches_k_color", value = c("skyblue", "darkorange", "grey","darkorange4","deepskyblue4",
                                    "darkolivegreen3","darkorchid1","firebrick","gold1","aquamarine2","brown2","darkseagreen3","black"),k = 13) %>%
  set("labels_cex", 0.7)  %>%
  set("leaves_pch", 19)  %>% 
  set("nodes_cex", 0.4) %>% 
  plot(axes=FALSE)

# Add the colored bar
colored_bars(colors = my_colors1, dend = hce, rowLabels = "subfam1")
legend("topright", legend=levels(subfam1), col= c("cyan", "green", "red", "darkblue", "yellow"), pch=15)

```

For the creation of the Dendrogramm only the first 17 PC scores were considered, which contribute to about 95.3% of variation. The Gelinae cluster over the different subfamilies with one bigger cluster on the right side. The fossil clusters between an Ichneumoninae and Pimplininae. It is on the right side, but still not in the one cluster of Gelinae at the very right end.
