---
title: "First_Analysis"
author: "Robin von Allmen"
date: "18 3 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
require(rgl)
options(rgl.useNULL = TRUE) # Suppress the separate window.
knitr::knit_hooks$set(webgl = hook_webgl)
library(geomorph)
setwd("~/GitHub/Ichneumonidae/R-Scripts/Determine Higher-Crouping")
```

## Preparing Data


```{r cars, include=FALSE}

landmark = readland.tps("blockcourse_highersub.tps",specID = "ID", readcurves = T, warnmsg = TRUE)

##      define classifiers (subfam or tribes)
species = as.factor( substr( dimnames(landmark)[[3]], 1, 6))
subfam = as.factor( substr( dimnames(landmark)[[3]], 8, 13))
areolet = as.factor( substr( dimnames(landmark)[[3]], 15, 15))
tribe = as.factor( substr( dimnames(landmark)[[3]], 17, 19))


##      define Semilandmarks, first LM is the start, last one the end (those are both
##      fixed landmarks, in between are the semilandmarks):
sliders_red = define.sliders(c(14,21:28,15))

##      procrustes fit - "superimposition" - rotates and scales the landmarks to their closest distance,
##      minimizing the shape and size difference:
proD<-gpagen(landmark[2:29,,], surfaces = NULL, curves= sliders_red,
             PrinAxes = FALSE, max.iter = NULL, ProcD = F, Proj = TRUE,
             print.progress = TRUE)
use= c("pimpli", "banchi", "gelina", "ophion", "ichneu")
reduced<- proD$coords[,,subfam%in% use]
pro_red<-gpagen(reduced, surfaces = NULL, curves= sliders_red,
             PrinAxes = FALSE, max.iter = NULL, ProcD = F, Proj = TRUE,
             print.progress = TRUE)
subfam1<-as.factor(substr(dimnames(pro_red$coords)[[3]],8,13))
species1<-as.factor(substr(dimnames(pro_red$coords)[[3]],1,6))
tribe1<-as.factor(substr(dimnames(pro_red$coords)[[3]],17,19))
```

## PCA Analysis
Here, A PCA  with the subfamilies Pimplinae, Gelinae, Ophioninae, Ichenumoninae and Banichnae without landmark 1 was done. 

```{r pressure, fig.show='hide', results='hide'}

PCA <-plotTangentSpace(pro_red$coords, groups=subfam1, legend=T)


```

One can work on this plot to get a beautiful shape so that it loocks somewhat like this:

```{r plot,echo=FALSE}
  ## the more beautiful version:

mat <- matrix(c(4,5,0,1,1,2,1,1,3), 3)
layout(mat, widths=c(1,1,1), heights=c(1,1,1))# set the size of the rows and columns

par(mar=c(4, 4, 1, 1)) # sets the margins
plot(PCA$pc.scores[,1],PCA$pc.scores[,2], col=subfam1, asp=T, las=1,pch=16, cex=1.5)
legend("topright", legend= levels(subfam1), col=palette(), pch=rep(19,14) )

segments(min(PCA$pc.scores[,1]), 0,  max(PCA$pc.scores[,1]),0, col="grey", lty="dashed")
segments(0, min(PCA$pc.scores[,2]), 0,  max(PCA$pc.scores[,2]), col="grey", lty="dashed")
# in order to name individuals: text(PCA$pc.scores[,1],PCA$pc.scores[,2], species)

ref <- mshape(pro_red$coords)
#qw= c(1,4,2,6,5,7,3)
par(mar = c(0,0,0,0)) # sets the margins

plotRefToTarget(ref,PCA$pc.shapes$PC1min)
# Item 3
plotRefToTarget(ref,PCA$pc.shapes$PC1max)
# Item 4
plotRefToTarget(ref,PCA$pc.shapes$PC2max)
# Item 5
plotRefToTarget(ref,PCA$pc.shapes$PC2min)
```


On the plot above we see, that Gelinae (Townes) are fairly good distinguishable from the other Tribes. The Data however does not represent modern taxonomy: Although most specimen still belong together in the same subfamily or Tribe, there has been a lot of improvement of building phenetic relationships between these genera together with molecular Data, hence some of the outliers in this plot that seem to belong in other groups, may actually be accounted to other taxanomic groups, as shown for the Genus Hemigaster for example. 



with the PCA scores u are also able to generate 3d plots which compare several Principal Components at once:

```{r 3D, webgl = TRUE}
colors = c("red","cyan","magenta","yellow","green")[subfam1]
plot3d(PCA$pc.scores[,1:3], col = colors, size = 9)
text3d(PCA$pc.scores[,1:3],texts=species1,pos=4,cex= 0.6)
legend3d("topright", legend = levels(subfam1), col = c("red","cyan","magenta","yellow","green") , pch= 16)
```

## Cluster Analysis
```{r Procrustes Distances, message = FALSE, results='hide', include=FALSE}
library(Evomorph)

inlist = list(pro_red$coords [,,])
for ( j in 1: dim(pro_red$coords )[3] )
  inlist[[j]] = pro_red$coords [,,j]


ID = paste (subfam1,species1, sep = "_", collapse = NULL, recycle0 = FALSE)

### PC1 for PC1dist thingy.. for empty matrix: ###################
PCA_sc<-PCA$pc.scores
row.names(PCA_sc) = ID
PCdist = as.matrix( dist(PCA_sc))

###  make empty matrix 
Proc_dist= matrix( NA, dim(PCA_sc)[1], dim(PCA_sc)[1] )

colnames(Proc_dist) = ID
rownames(Proc_dist) = ID

###  make a loop : calculate proc distances for each row. 
Proc_dist
for ( i in 1:length(ID)){
  Proc_dist[i,] = ShapeDist(inlist, inlist[[i]])
}

write.table(Proc_dist,"Proc_distances_ban_pim_oph_cte.txt", sep=" ",row.names = TRUE, col.names = TRUE)

###################################################################
Proc_dist = read.table("Proc_distances_ban_pim_oph_cte.txt", header=T)
```

```{r Clusters,}

#Data of Alex does not include tribe
#old CA with PCA values, now conducting with Procrustes distances
#data = data.frame(species1, subfam1, PCA$pc.scores[,1:6])

# Compute distances and hierarchical clustering
dd <- dist(scale(Proc_dist))

## Ward's minimum variance method
hc <- hclust(dd, method = "complete")

hc$labels = row.names(Proc_dist)

# Convert hclust into a dendrogram and plot
hcd <- as.dendrogram(hc)
```
## Dendrogramm
```{r Bäumli, results=FALSE}
require("dendextend") ## needs a dendrogram


# Create a vector of colors, darkgreen if am is 0, green if 1.
my_colors <- ifelse(substr(hc$labels,1,6)=="banchi", "cyan",ifelse(substr(hc$labels,1,6)=="gelina","green", ifelse(substr(hc$labels,1,6)=="ichneu", "red",ifelse(substr(hc$labels,1,6)=="ophion", "darkblue",ifelse(substr(hc$labels,8,13)== "fossil", "magenta", "yellow" )))))
#my_colors <- ifelse(hc$labels=="scolob", "forestgreen",ifelse(hc$labels=="metopi", "green","yellow"))

par(mar=c(10,5,1,1))
hcd %>%
  set("labels_col", value = c("skyblue", "darkorange", "grey","darkorange4","deepskyblue4",
                              "darkolivegreen3","darkorchid1","firebrick","gold1","aquamarine2","brown2","darkseagreen3","black"), k=13) %>%
  set("branches_k_color", value = c("skyblue", "darkorange", "grey","darkorange4","deepskyblue4",
                                    "darkolivegreen3","darkorchid1","firebrick","gold1","aquamarine2","brown2","darkseagreen3","black"),k = 13) %>%
  set("labels_cex", 0.8)  %>%
  set("leaves_pch", 19)  %>% 
  set("nodes_cex", 0.4) %>%
  plot(axes=FALSE)

# Add the colored bar
colored_bars(colors = my_colors, dend = hcd, rowLabels = "subfamily")
legend("topright", legend=levels(subfam1), col= c("cyan","magenta", "green", "red", "darkblue", "yellow"), pch=15)

```

For the creation of the Dendrogramm only the first 6 PC scores were considered, which contribute to about 82.5% of variation. Higher PC scores did not seem to reflect the data that would be observed on the PCA plots, as the fossil was usually clustered along with a specimen from Echthrini that in contrast was also clustered in Gelinae eventhough the fossil is much more distant from Echthrini than Gelinae.
