---
title: "Gelinae_vs_Ichneumoninae_Analysis"
author: "Robin von Allmen"
date: "18 3 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
require(rgl)
options(rgl.useNULL = TRUE) # Suppress the separate window.
knitr::knit_hooks$set(webgl = hook_webgl)
library(geomorph)
setwd("~/GitHub/Ichneumonidae/R-Scripts")
```

## Preparing Data


```{r cars, results=FALSE}
setwd("~/GitHub/Ichneumonidae/R-Scripts")
landmark = readland.tps("DATA_GEL_ICH.tps",specID = "ID", readcurves = T, warnmsg = TRUE)
landmark2 = readland.tps("DATA_GEL_ICH.tps",specID = "ID", readcurves = F, warnmsg = TRUE)
##      define classifiers (subfam or tribes)
species = as.factor( substr( dimnames(landmark)[[3]], 1, 6))
subfam = as.factor( substr( dimnames(landmark)[[3]], 8, 13))
areolet = as.factor( substr( dimnames(landmark)[[3]], 15, 15))
tribe = as.factor( substr( dimnames(landmark)[[3]], 17, 19))


##      define Semilandmarks, first LM is the start, last one the end (those are both
##      fixed landmarks, in between are the semilandmarks):
sliders = define.sliders(c(15,22:29,16))

##      procrustes fit - "superimposition" - rotates and scales the landmarks to their closest distance,
##      minimizing the shape and size difference:
proD<-gpagen(landmark, surfaces = NULL, curves = sliders,
             PrinAxes = FALSE, max.iter = NULL, ProcD = F, Proj = TRUE,
             print.progress = TRUE)

proD2<-gpagen(landmark2, surfaces = NULL,
             PrinAxes = FALSE, max.iter = NULL, ProcD = F, Proj = TRUE,
             print.progress = TRUE)

```

## PCA Analysis

When comparing the Procrustes fit on the model that compares Ichneumoninae and Gelinae with eachother, we are surprised by the fact, that the curve between LM 15-16 does not improve dinstiguishment of the two subfamilies. Hence we proceed in this case with the analyses only obtained from fixed Landmarks.

```{r plot1}

PCA <-plotTangentSpace(proD$coords, groups=subfam, legend=T)
legend("topright", legend= levels(tribe), col=palette(), pch=rep(19,14) )
title(main = "Gelinae vs Ichneumoninae - With Curves")


```

```{r plot2}
PCA1 <-plotTangentSpace(proD2$coords, groups=subfam, legend=T)
legend("topright", legend= levels(tribe), col=palette(), pch=rep(19,14) )
title(main = "Gelinae vs Ichneumoninae - Without Curves")

```

```{r plot,echo=FALSE}
  ## the more beautiful version:

mat <- matrix(c(4,5,0,1,1,2,1,1,3), 3)
layout(mat, widths=c(1,1,1), heights=c(1,1,1))# set the size of the rows and columns

par(mar=c(4, 4, 1, 1)) # sets the margins
plot(PCA1$pc.scores[,1],PCA1$pc.scores[,2], col=subfam, asp=T, las=1,pch=16, cex=1.5)
legend("topright", legend= levels(subfam), col=palette(), pch=rep(19,14) )

segments(min(PCA1$pc.scores[,1]), 0,  max(PCA1$pc.scores[,1]),0, col="grey", lty="dashed")
segments(0, min(PCA1$pc.scores[,2]), 0,  max(PCA1$pc.scores[,2]), col="grey", lty="dashed")
# in order to name individuals: text(PCA$pc.scores[,1],PCA$pc.scores[,2], species)

ref <- mshape(proD2$coords)
#qw= c(1,4,2,6,5,7,3)
par(mar = c(0,0,0,0)) # sets the margins

plotRefToTarget(ref,PCA1$pc.shapes$PC1min)
# Item 3
plotRefToTarget(ref,PCA1$pc.shapes$PC1max)
# Item 4
plotRefToTarget(ref,PCA1$pc.shapes$PC2max)
# Item 5
plotRefToTarget(ref,PCA1$pc.shapes$PC2min)
```


## 3D-Plot

with the PCA scores u are also able to generate 3d plots which compare several Principal Components at once:

```{r 3D, webgl = TRUE}
colors = c("red","cyan")[subfam]
plot3d(PCA1$pc.scores[,1:3], col = colors, size = 9)
text3d(PCA1$pc.scores[,1:3],texts=species,pos=4,cex= 0.6)
legend3d("topright", legend = c("Gelinae","Ichneumoninae"), col = c("red","cyan") , pch= 16)


```