---
title: "First_Analysis"
author: "Robin von Allmen"
date: "18 3 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
require(rgl)
options(rgl.useNULL = TRUE) # Suppress the separate window.
knitr::knit_hooks$set(webgl = hook_webgl)
library(geomorph)
setwd("~/GitHub/Ichneumonidae/R-Scripts")
```

## Preparing Data


```{r cars, include=FALSE}
setwd("~/GitHub/Ichneumonidae/R-Scripts")
landmark = readland.tps("DATA_MODERN_TAX.tps",specID = "ID", readcurves = T, warnmsg = TRUE)

##      define classifiers (subfam or tribes)
species = as.factor( substr( dimnames(landmark)[[3]], 1, 6))
subfam = as.factor( substr( dimnames(landmark)[[3]], 8, 13))
areolet = as.factor( substr( dimnames(landmark)[[3]], 15, 15))
tribe = as.factor( substr( dimnames(landmark)[[3]], 17, 19))


##      define Semilandmarks, first LM is the start, last one the end (those are both
##      fixed landmarks, in between are the semilandmarks):
sliders = define.sliders(c(15,22:29,16))

##      procrustes fit - "superimposition" - rotates and scales the landmarks to their closest distance,
##      minimizing the shape and size difference:
proD<-gpagen(landmark, surfaces = NULL, curves= sliders,
             PrinAxes = FALSE, max.iter = NULL, ProcD = F, Proj = TRUE,
             print.progress = TRUE)
```

## PCA Analysis


```{r pressure}
PCA <-plotTangentSpace(proD$coords, groups=subfam, legend=T)
#PCA$pc.summary
dev.off()

#one can look at different PCA combinations to find the best model for its purpose
plot(PCA$pc.scores[,1],PCA$pc.scores[,2], col=subfam, asp=T, las=1,pch=16, cex=1.3)
legend("topright", legend= levels(subfam), col=palette(), pch=rep(19,14) )

# use the text() function to get the names of individual points
#text(PCA$pc.scores[,1],PCA$pc.scores[,2],labels=species,cex=0.7,pos=4)

dev.off()
```
One can work on this plot to get a beautiful shape so that it loocks somewhat like this:

```{r plot,echo=FALSE}
  ## the more beautiful version:

mat <- matrix(c(4,5,0,1,1,2,1,1,3), 3)
layout(mat, widths=c(1,1,1), heights=c(1,1,1))# set the size of the rows and columns

par(mar=c(4, 4, 1, 1)) # sets the margins
plot(PCA$pc.scores[,1],PCA$pc.scores[,2], col=subfam, asp=T, las=1,pch=16, cex=1.5)
legend("topright", legend= levels(subfam), col=palette(), pch=rep(19,14) )

segments(min(PCA$pc.scores[,1]), 0,  max(PCA$pc.scores[,1]),0, col="grey", lty="dashed")
segments(0, min(PCA$pc.scores[,2]), 0,  max(PCA$pc.scores[,2]), col="grey", lty="dashed")
# in order to name individuals: text(PCA$pc.scores[,1],PCA$pc.scores[,2], species)

ref <- mshape(proD$coords)
#qw= c(1,4,2,6,5,7,3)
par(mar = c(0,0,0,0)) # sets the margins

plotRefToTarget(ref,PCA$pc.shapes$PC1min)
# Item 3
plotRefToTarget(ref,PCA$pc.shapes$PC1max)
# Item 4
plotRefToTarget(ref,PCA$pc.shapes$PC2max)
# Item 5
plotRefToTarget(ref,PCA$pc.shapes$PC2min)
```


On the plot above we see, that Gelinae (Townes) are fairly good distinguishable from the other Tribes. The Data however does not represent modern taxonomy: Although most specimen still belong together in the same subfamily or Tribe, there has been a lot of improvement of building phenetic relationships between these genera together with molecular Data, hence some of the outliers in this plot that seem to belong in other groups, may actually be accounted to other taxanomic groups, as shown for the Genus Hemigaster for example. 



with the PCA scores u are also able to generate 3d plots which compare several Principal Components at once:

```{r 3D, webgl = TRUE}
colors = c("red","cyan","magenta","yellow")[subfam]
plot3d(PCA$pc.scores[,1:3], col = colors, size = 9)
text3d(PCA$pc.scores[,1:3],texts=species,pos=4,cex= 0.6)
legend3d("topright", legend = c("Ateleute","Cryptinae", "Gelinae", "Phygadeutinae"), col = c("red","cyan","magenta","yellow") , pch= 16)
```
